\section{Design Goals}

PMChecker aims at verifying a concurrent persistent memory data structure
for crash consistency and correctness.

What is correctness?
A persistent memory data structure is correct if all the safety critical
operations are ordered using memory fence instructions.

In detail, Persistent Memory controller reorders load and store instructions
to optimize PM accesses. If a data structure requests a memory allocation
and then writes a persistent memory pointer to the allocated memory,
unless the data structure orders these operations with fences,
the PM controller can reorder these operations, resulting in a corrupted
state of the data structure.

Re-ordering two instructions is okay if they operate idependently.
Dependent operations might result in data corruption if reordered.
PMChecker should check if the operations without fences are commutative.
If the operations are commutative then there is no data corruption.
But if not, then there is a potential race condition.
What are the minimal set of re-orderings that PMChecker should consider?
How can PMChecker eliminate false positives.

PMChecker is designed to have the following goals.
\begin{enumerate}
	\item Little to no false positives
	\item Should work with diverse implementations of 
		persistent memory data structures
	\item Should be easy to use similar to push-button verification,
		with no or minimum developer involvement.
	\item Should take advantage of the existing black-box and
		white-box verification techniques.
\end{enumerate}

1) Correctness: Un ordered operations should be commutative
Linearization points should be strictly ordered.
Operations between linearization points should be commutative.
2) Crash consistency: Isolation translates to crash consistency.
3) Crash recovery: Operations should be idempotent in the recovery


Linearization points: How to identify linearization points?
Reordering capsules: What all reorderings to check?
Commutative capsules
Idempotent capsules

How to construct a program graph and use the decades of research
along the lines of model checking to construct the state space
of PMChecker? 
