\section{WORT}

This section describes WORT~\cite{201600}, write-optimized
radix tree. WORT is used in the rest of this work to describe
the modeling of PM behavior and to verifying consistency properties.

\vheading{Radix tree}. Several B-tree based indexing structures
have been proposed for PM. These indexes focused on reducing the
number of calls to the expensive memory fence and cache line flush
instructions. They employ an append-only update strategy. Such
strategies have been shown to significantly reduce duplicate copies
needed for schemes such as logging resulting in improved performance.
However, this strategy does not allow these indexes to retain one of
the key features of B-trees, that is, having the keys sorted. Moreover,
this strategy is insufficient in handling node overflows as node splits
involve multiple node changes, making logging necessary.

The first contribution of WORT is showing the appropriateness and the
limitation of radix trees for PM. That is, since the radix tree structure
is determined by the prefix of the inserted keys, the radix tree does not
require key comparisons. Furthermore, tree rebalancing operations and
updates in node granularity units are also not necessary. Instead,
insertion or deletion of a key results in a single 8-byte update operation,
which is perfect for PM. However, the traditional radix tree is known
to poorly utilize memory and cache space. In order to overcome this limitation,
the radix tree employs a path compression optimization, which combines
multiple tree nodes that form a unique search path into a single node.
Although path compression significantly improves the performance of the
radix tree, it involves node split and merge operations,
which is detrimental for PM.

\vheading{Write optimized radix tree}. The second contribution of WORT is
the desing of an efficient write-optimized, path compression algorithm for
radix trees. In this paper, we model and verify the path-compressed and
write-optimized radix tree for PM.